---
title: "Trading Strategy: TMA in Metal Commodities Markets"
subtitle: "CA01 - Investment Fundamentals, Group Project"
author: "Group 8 - Benedikt Jaletzke"
date: "12/10/2020"
abstract: "Triangular Moving Averages in Metal Commodities Markets for Consistent Long-Only Returns"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
PLEASE USE THIS VERSION
# Official Checklist:
- Introduction
- Economic Rationale
- Data (Sources)
- Method Used
  - Technical Indicator
  - Market
- Main Findings
- Appendix 


# Introduction

The idea for this group project was to come up with a strategy that either expoited an arbitrage opportunity in the market, or represents another form of hedge fund strategy. Our group decided to focus on using technical indicators. Within this area, we specifically chose moving averages, and as a suitable market for their application settled on metal commodities.


# Economic Rationale


# Method Used

## Portfolio Allocation

In order to ideally work within the metals markets, the portfolio will consist ocmpletely of two already existing indices that are calculated on a daily basis, the S&P Precious Metals Index, consisting of [TBC] and the S&P Industrial Metals, containing [TBC]. By using these two indices as our baseline instead of attempting to construct our own, we can create a two-security portfolio with different weights to find the best possible results. As for the choice of allocations, we tried different combination weights of the indices. 

To construct the efficient portfolio, we had to calculate the return and variance of teh two separate indices. 
```{r}
# Return Variance Calculation
# Combining both indices into one frame, cleaning it up a little
combined_indices <- full_join(industrial_1, precious_1, by = "date") %>% 
  rename(precious_metals = index.y,
         pm_close = close.y,
         ind_close = close.x,
         industrial_metals = index.x) %>% 
  subset(select = -c(ticker.x, ticker.y)) %>%
  head(-2)
  

#return and std
ind_price <- na.approx(zoo(combined_indices$ind_close, combined_indices$date))
prec_price <- na.approx(zoo(combined_indices$pm_close, combined_indices$date))

rets_industry <- Return.calculate(xts(ind_price),  method = "discrete")
names(rets_industry) <- "return_ind"
rets_prec <- Return.calculate(xts(prec_price),  method = "discrete")
names(rets_prec) <- "return_prec"

combined_indices <- cbind(combined_indices, rets_industry, rets_prec)

stats <- combined_indices %>% 
  summarise(mean_ind = mean(return_ind[-1]), mean_prec = mean(return_prec[-1]),
            sd_ind = sd(return_ind[-1]), sd_prec = sd(return_prec[-1]))
```


```{r}
# 50/50 Portfolio
w_ind <- 0.5
w_prec <- 1-w_ind

equal_portfolio <-  precious_metals * w_prec + industrial_metals * w_ind
```

```{r}
# Efficient Portfolio
stats <- combined_indices %>% 
  summarise(mean_ind = mean(return_ind[-1]), mean_prec = mean(return_prec[-1]),
            sd_ind = sd(return_ind[-1]), sd_prec = sd(return_prec[-1]),
            corr_ind = cor(return_ind[-1],return_prec[-1]))

sh_r_max <- function(x){
  (x*stats$mean_ind+(1-x)stats$mean_prec)/sqrt(x^2*stats$sd_ind^2+(1-x)^2*stats$sd_prec^2+2*x(1-x)*stats$sd_ind*stats$sd_prec*stats$corr_ind)
}

opt_port <- optimize(f = sh_r_max, lower = -1,upper = 1, maximum = TRUE)

weight_ind <- opt_port$objective
```

```{r}
# Combination Graph
```


## Technical Indicator

From https://uc-r.github.io/ts_moving_averages

The concept of simple moving averages can be extended to taking moving averages of moving averages. This technique is often employed with an even number of data points so that the final product is symmetric around each point. 
An even-numbered moving average is unbalanced, and for our purposes, the unbalancing will be in favor of more recent observations. For example, to calculate a 4-MA, the equation is as follows:
^yt=(yt−1+yt+yt+1+yt+2)/4

To make the moving average symmetric (and therefore more accurate), we then take a 2-MA of the 4-MA to create a 2 x 4-MA. For the 2-MA step, we average the current and previous moving averages, thus resulting in an overall estimate of:

```{r}
# rolling averages. Useful for later calculations
# k = number of days used in average
metals_tma <- combined_data %>% 
  mutate(index_ma01 = rollmean(index, k=1, fill=NA),
         index_ma05 = rollmean(index, k=90, fill=NA))
```

### Finding appropriate Modifiers for the TMA

The next challenge lies in finding the appropriate time periods for the TMA. As it is an average of averages, we have to consider which of these will give us the best return.

## Market Choice
(Maybe show correlation between all metals? Simple GGPairs)

### Correlation Graph for the Choice Metals
```{r}
ggcorr(combined_indices) + labs(
  title = "Correlation between Metal Prices")

ggpairs(wind)
```

# Results

## Charts

## Data


# Appendix
## Libraries
```{r}
library(lubridate)
library(tidyverse)
library(GGally)
library(skimr)
library(lubridate)
library(fpp2) 
library(zoo)
library(PerformanceAnalytics)
```


## Importing Indices - Metals
```{r}
# Data from Finaeon
precious <- read.csv("data/sp_precmet.csv", skip = 2)
industrial <- read.csv("data/sp_indmet.csv", skip = 2)

# S&P Precious Metals Index
precious_1 <- precious %>%
  mutate(date = as.Date(precious$Date, format = "%m/%d/%Y")) %>% 
  filter(date > as.Date("1989-12-31")) %>% 
  subset(data = precious, select = -c(Date, Open, High, Low)) %>% 
  rename(close = Close, 
         ticker = Ticker) %>% 
  mutate(index = close/close[1])

# S&P Industrial Metals Index
industrial_1 <- industrial %>%
  mutate(date = as.Date(industrial$Date, format = "%m/%d/%Y")) %>% 
  filter(date > as.Date("1989-12-31")) %>% 
  subset(data = industrial, select = -c(Date, Open, High, Low)) %>% 
  rename(close = Close, 
         ticker = Ticker) %>% 
  mutate(index = close/close[1])

# Combining both indices into one frame, cleaning it up a little
combined_indices <- full_join(industrial_1, precious_1, by = "date") %>% 
  rename(precious_metals = index.y,
         pm_close = close.y,
         ind_close = close.x,
         industrial_metals = index.x) %>% 
  subset(select = -c(ticker.x, ticker.y))

# Simple Plot
ggplot() + 
  geom_line(data = combined_indices, aes(x=date, y=precious_metals), color = "red") + 
  geom_line(data = combined_indices, aes(x=date, y=industrial_metals), color = "green")

# Weights
eq_w_pm <- 0.5
eq_w_ind <- 1-eq_w_pm

# calculate index
wind <- combined_indices %>% 
  mutate(eq_ind = w_pm * precious_metals + w_ind * industrial_metals) %>% 
  subset(select = c(date, eq_ind))
# Index plot vs. Separate Lines
ggplot() + 
  geom_line(data = wind, aes(x=date, y=eq_ind), color="red")+ 
  geom_line(data = combined_indices, aes(x=date, y=precious_metals), color = "blue") + 
  geom_line(data = combined_indices, aes(x=date, y=industrial_metals), color = "green")
```

## Importing and Filtering Equities Indices for Comparison Purposes
```{r}
# Importing Index Data
# Filtering: Dates, Unused Indices (keeping )
global_indices <- read.csv("data/global.csv", skip = 51)

global_indices_filtered <- global_indices %>%
  mutate(date = as.Date(global_indices$Date, format = "%Y/%m/%d")) %>% 
  filter(date > as.Date("1989-12-31")) %>% 
  subset(data = global_indices, select = c(Date, X_NDXD_Close, X_DJA3D_Close, X_GDAXIPD_Close, X_SPXD_Close))
```


# Old Stuff
## OLD - Equal Weighted Separate Metals
```{r}
# Equally Weighted Portfolio of Metals
w_pal <- 0.2
w_plat <- 0.2
w_ni <- 0.2
w_cop <- 0.2
w_tin <- 0.2

weighted_index = sep_ind_data %>% 
  mutate(big5_index = (w_pal * palladium_ind + w_plat * platinum_ind + w_ni * nickel_ind + w_cop * copper_ind)) %>% 
  subset(select = c(date, big5_index))
  
ggplot()+geom_line(data = combined_data, aes(x=date, y=index, color=ticker)) + geom_line(data = weighted_index, aes(x=date, y=big5_index)) +geom_line() +scale_y_log10()

```

## OLD - Overview Graph
### Development of the Metal prices over time (indexed).
```{r, fig.width = 11}
#Basic Index/Time Graph
ggplot(combined_data, aes(x=date, y=index, color=ticker)) + 
  geom_line() +
  scale_color_discrete(
    name = "Metal",
    labels = c("Copper", "Tin", "Platinum", "Palladium", "S&P Index", "Nickel")) +
  theme_minimal() +
  scale_y_log10() +
  labs(
    title = "Metal Commodities Price Development",
    subtitle = "Indexed Jan 1st, 1989",
    x = "Date", 
    y = "Index",
    caption = "Source: Finaeon")

unique(combined_data$ticker)
```




## OLD - Importing Metal Commodities Prices
```{r}
copper <- read.csv("data/copper.csv", skip = 2)
tin <- read.csv("data/tin.csv", skip = 2)
platinum <- read.csv("data/platinum.csv", skip = 2)
palladium <- read.csv("data/palladium.csv", skip = 2)
nickel <- read.csv("data/nickel.csv", skip = 2)
```

### Data Cleaning of the Metals Returns
```{r}
# Transforming Data to be more Usable for this analysis: 
# Filtering Dates, unused columns, cleaning up names

copper_1 <- copper %>%
  mutate(date = as.Date(copper$Date, format = "%m/%d/%Y")) %>% 
  filter(date > as.Date("1989-12-31")) %>% 
  subset(data = copper, select = -c(Date, Open, High, Low)) %>% 
  rename(close = Close, 
         ticker = Ticker) %>% 
  mutate(index = close/close[1])

tin_1 <- tin %>%
  mutate(date = as.Date(tin$Date, format = "%m/%d/%Y")) %>% 
  filter(date > as.Date("1989-12-31")) %>% 
  subset(data = tin, select = -c(Date, Open, High, Low)) %>% 
  rename(close = Close, 
         ticker = Ticker) %>% 
  mutate(index = close/close[1])

platinum_1 <- platinum %>%
  mutate(date = as.Date(platinum$Date, format = "%m/%d/%Y")) %>% 
  filter(date > as.Date("1989-12-31")) %>%  
  subset(data = platinum, select = -c(Date, Open, High, Low)) %>% 
  rename(close = Close, 
         ticker = Ticker) %>% 
  mutate(index = close/close[1])

palladium_1 <- palladium %>%
  mutate(date = as.Date(palladium$Date, format = "%m/%d/%Y")) %>% 
  filter(date > as.Date("1989-12-31")) %>% 
  subset(data = palladium, select = -c(Date, Open, High, Low)) %>% 
  rename(close = Close, 
         ticker = Ticker) %>% 
  mutate(index = close/close[1])

nickel_1 <- nickel %>%
  mutate(date = as.Date(nickel$Date, format = "%m/%d/%Y")) %>% 
  filter(date > as.Date("1989-12-31")) %>% 
  subset(data = nickel, select = -c(Date, Open, High, Low)) %>% 
  rename(close = Close, 
         ticker = Ticker) %>% 
  mutate(index = close/close[1])





# Combining Data into one data frame

p1 <- full_join(copper_1, tin_1, by = c("date", "ticker", "close", "index"))
p2 <- full_join(p1, platinum_1, by = c("date", "ticker", "close", "index"))
p3 <- full_join(p2, palladium_1, by = c("date", "ticker", "close", "index"))
p4 <- full_join(p3, precious_1, by = c("date", "ticker", "close", "index"))
combined_data <- full_join(p4, nickel_1, by = c("date", "ticker", "close", "index"))

#One data frame, separate indices
p1_2 <- left_join(copper_1, tin_1, by = "date")
p2_2 <- left_join(p1_2, platinum_1, by = "date")
p3_2 <- left_join(p2_2, palladium_1, by = "date")
sep_ind_data <- left_join(p3_2, nickel_1, by = "date")

sep_ind_data <-  sep_ind_data %>% 
  rename(copper_ind=index.x , 
         tin_ind=index.y,
         platinum_ind=index.x.x,
         palladium_ind=index.y.y,
         nickel_ind=index)

sep_ind_data <- sep_ind_data %>% 
 subset(data = sep_ind_data, select = -c(ticker.x, ticker.y, ticker.y.y, ticker.x.x))
sep_ind_data <- sep_ind_data %>% 
  subset(data = sep_ind_data, select = -c(close.x, close.y, close.y.y, close.x.x))
sep_ind_data <- sep_ind_data %>% 
  subset(data = sep_ind_data, select = -c(ticker, close))
```
