---
title: "Trading Strategy: TMA in Metal Commodities Markets"
subtitle: "CA01 - Investment Fundamentals, Group Project"
author: "Group 8 - Benedikt Jaletzke"
date: "12/10/2020"
abstract: "Triangular Moving Averages in Metal Commodities Markets for Consistent Long-Only Returns"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
DEPRECEATED!
Checklist:
- Introduction
- Economic Rationale
- Data (Sources)
- Method Used
  - Technical Indicator
  - Market
- Main Findings

# Introduction

The idea for this group project was to come up with a strategy that either expoited an arbitrage opportunity in the market, or represents another form of hedge fund strategy. Our group decided to focus on using technical indicators. Within this area, we specifically chose moving averages, and as a suitable market for their application settled on metal commodities.

# Economic Rationale

# Market Choice - Why Metals?

Metal commodities in general display a trend that is somewhat sought after, which is that they generally move in unison.

Correlation Graph for the Choice Metals
```{r}
ggcorr(sep_ind_data) + labs(
  title = "Correlation between Metal Prices"
)
ggpairs(sep_ind_data)
```

Development of the Metal prices over time (indexed).
```{r, fig.width = 11}
#Basic Index/Time Graph
ggplot(combined_data, aes(x=date, y=index, color=ticker)) + 
  geom_line() +
  scale_color_discrete(
    name = "Metal",
    labels = c("Copper", "Tin", "Platinum", "Palladium", "S&P Index", "Nickel")) +
  theme_minimal() +
  scale_y_log10() +
  labs(
    title = "Metal Commodities Price Development",
    subtitle = "Indexed Jan 1st, 1989",
    x = "Date", 
    y = "Index",
    caption = "Source: Finaeon")

unique(combined_data$ticker)
```

# Choice of Technical Indicator

From https://uc-r.github.io/ts_moving_averages

The concept of simple moving averages can be extended to taking moving averages of moving averages. This technique is often employed with an even number of data points so that the final product is symmetric around each point. 
An even-numbered moving average is unbalanced, and for our purposes, the unbalancing will be in favor of more recent observations. For example, to calculate a 4-MA, the equation is as follows:
^yt=(yt−1+yt+yt+1+yt+2)/4

To make the moving average symmetric (and therefore more accurate), we then take a 2-MA of the 4-MA to create a 2 x 4-MA. For the 2-MA step, we average the current and previous moving averages, thus resulting in an overall estimate of:

```{r}
# rolling averages. Useful for later calculations
# k = number of days used in average
metals_tma <- combined_data %>% 
  mutate(index_ma01 = rollmean(index, k=1, fill=NA),
         index_ma05 = rollmean(index, k=90, fill=NA))
```

# Implementation

## Constructing an Index

Using the formula for weights of a >2 asset portfolio, we can construct an index of the five chosen metals that we can the apply the indicator to. 
In order to do this, we first have to compute the annual returns, correlation and standard deviation for the metals.
```{r}
# E(r)
er_assets <- combined_data %>% 
  group_by(ticker) %>% 
  Return.annualized(scale = 12, geometric = FALSE)
# StDev

# Correlation

```
Equal Weighted Separate Metals
```{r}
# Equally Weighted Portfolio of Metals
w_pal <- 0.2
w_plat <- 0.2
w_ni <- 0.2
w_cop <- 0.2
w_tin <- 0.2

weighted_index = sep_ind_data %>% 
  mutate(big5_index = (w_pal * palladium_ind + w_plat * platinum_ind + w_ni * nickel_ind + w_cop * copper_ind)) %>% 
  subset(select = c(date, big5_index))
  
ggplot()+geom_line(data = combined_data, aes(x=date, y=index, color=ticker)) + geom_line(data = weighted_index, aes(x=date, y=big5_index)) +geom_line() +scale_y_log10()

```

## Finding appropriate Modifiers for the TMA
The next challenge lies in finding the appropriate time periods for the TMA. As it is an average of averages, we have to consider which of these will give us the best return.



## Trading Modifiers for the TMA


# Results

## Charts

## Data

# Appendix

Libraries
```{r}
library(lubridate)
library(tidyverse)
library(GGally)
library(skimr)
library(lubridate)
library(fpp2) 
library(zoo)
library(PerformanceAnalytics)
```

Importing Indices - Metals
```{r}
# Data from Finaeon
precious <- read.csv("data/sp_precmet.csv", skip = 2)
industrial <- read.csv("data/sp_indmet.csv", skip = 2)

# S&P Precious Metals Index
precious_1 <- precious %>%
  mutate(date = as.Date(precious$Date, format = "%m/%d/%Y")) %>% 
  filter(date > as.Date("1989-12-31")) %>% 
  subset(data = precious, select = -c(Date, Open, High, Low)) %>% 
  rename(close = Close, 
         ticker = Ticker) %>% 
  mutate(index = close/close[1])

# S&P Industrial Metals Index
industrial_1 <- industrial %>%
  mutate(date = as.Date(industrial$Date, format = "%m/%d/%Y")) %>% 
  filter(date > as.Date("1989-12-31")) %>% 
  subset(data = industrial, select = -c(Date, Open, High, Low)) %>% 
  rename(close = Close, 
         ticker = Ticker) %>% 
  mutate(index = close/close[1])

combined_indices <- full_join(industrial_1, precious_1, by = "date")

ggplot() + 
  geom_line(data = combined_indices, aes(x=date, y=index.x, color="red")) + 
  geom_line(data = combined_indices, aes(x=date, y=index.y, color="green"))
```



Importing Metal Commodities Prices
```{r}
copper <- read.csv("data/copper.csv", skip = 2)
tin <- read.csv("data/tin.csv", skip = 2)
platinum <- read.csv("data/platinum.csv", skip = 2)
palladium <- read.csv("data/palladium.csv", skip = 2)
nickel <- read.csv("data/nickel.csv", skip = 2)

# nickel_new <- read.csv("data/nickel_new.csv", skip = 2)
```

Importing and Filtering Indices for Comparison Purposes
```{r}
# Importing Index Data
# Filtering: Dates, Unused Indices (keeping )
global_indices <- read.csv("data/global.csv", skip = 51)
global_indices_filtered <- global_indices %>%
  mutate(date = as.Date(global_indices$Date, format = "%Y/%m/%d")) %>% 
  filter(date > as.Date("1989-12-31")) %>% 
  subset(data = global_indices, select = c(Date, X_NDXD_Close, X_DJA3D_Close, X_GDAXIPD_Close, X_SPXD_Close))
```

Data Cleaning of the Metals Returns
```{r}
# Transforming Data to be more Usable for this analysis: 
# Filtering Dates, unused columns, cleaning up names

copper_1 <- copper %>%
  mutate(date = as.Date(copper$Date, format = "%m/%d/%Y")) %>% 
  filter(date > as.Date("1989-12-31")) %>% 
  subset(data = copper, select = -c(Date, Open, High, Low)) %>% 
  rename(close = Close, 
         ticker = Ticker) %>% 
  mutate(index = close/close[1])

tin_1 <- tin %>%
  mutate(date = as.Date(tin$Date, format = "%m/%d/%Y")) %>% 
  filter(date > as.Date("1989-12-31")) %>% 
  subset(data = tin, select = -c(Date, Open, High, Low)) %>% 
  rename(close = Close, 
         ticker = Ticker) %>% 
  mutate(index = close/close[1])

platinum_1 <- platinum %>%
  mutate(date = as.Date(platinum$Date, format = "%m/%d/%Y")) %>% 
  filter(date > as.Date("1989-12-31")) %>%  
  subset(data = platinum, select = -c(Date, Open, High, Low)) %>% 
  rename(close = Close, 
         ticker = Ticker) %>% 
  mutate(index = close/close[1])

palladium_1 <- palladium %>%
  mutate(date = as.Date(palladium$Date, format = "%m/%d/%Y")) %>% 
  filter(date > as.Date("1989-12-31")) %>% 
  subset(data = palladium, select = -c(Date, Open, High, Low)) %>% 
  rename(close = Close, 
         ticker = Ticker) %>% 
  mutate(index = close/close[1])

nickel_1 <- nickel %>%
  mutate(date = as.Date(nickel$Date, format = "%m/%d/%Y")) %>% 
  filter(date > as.Date("1989-12-31")) %>% 
  subset(data = nickel, select = -c(Date, Open, High, Low)) %>% 
  rename(close = Close, 
         ticker = Ticker) %>% 
  mutate(index = close/close[1])





# Combining Data into one data frame

p1 <- full_join(copper_1, tin_1, by = c("date", "ticker", "close", "index"))
p2 <- full_join(p1, platinum_1, by = c("date", "ticker", "close", "index"))
p3 <- full_join(p2, palladium_1, by = c("date", "ticker", "close", "index"))
p4 <- full_join(p3, precious_1, by = c("date", "ticker", "close", "index"))
combined_data <- full_join(p4, nickel_1, by = c("date", "ticker", "close", "index"))

#One data frame, separate indices
p1_2 <- left_join(copper_1, tin_1, by = "date")
p2_2 <- left_join(p1_2, platinum_1, by = "date")
p3_2 <- left_join(p2_2, palladium_1, by = "date")
sep_ind_data <- left_join(p3_2, nickel_1, by = "date")

sep_ind_data <-  sep_ind_data %>% 
  rename(copper_ind=index.x , 
         tin_ind=index.y,
         platinum_ind=index.x.x,
         palladium_ind=index.y.y,
         nickel_ind=index)

sep_ind_data <- sep_ind_data %>% 
 subset(data = sep_ind_data, select = -c(ticker.x, ticker.y, ticker.y.y, ticker.x.x))
sep_ind_data <- sep_ind_data %>% 
  subset(data = sep_ind_data, select = -c(close.x, close.y, close.y.y, close.x.x))
sep_ind_data <- sep_ind_data %>% 
  subset(data = sep_ind_data, select = -c(ticker, close))
```


